---
title: "Introduction to sgejobs"
author: 
  - name: Leonardo Collado-Torres
    affiliation:
    - &libd Lieber Institute for Brain Development, Johns Hopkins Medical Campus
    - &ccb Center for Computational Biology, Johns Hopkins University
    email: lcolladotor@gmail.com
date: "`r doc_date()`"
package: "`r pkg_ver('sgejobs')`"
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Introduction to sgejobs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r vignetteSetup, echo=FALSE, message=FALSE, warning = FALSE}
## Track time spent on making the vignette
startTime <- Sys.time()

## Bib setup
library('knitcitations')

## Load knitcitations with a clean bibliography
cleanbib()
cite_options(hyperlink = 'to.doc', citation_format = 'text', style = 'html')
# Note links won't show for now due to the following issue
# https://github.com/cboettig/knitcitations/issues/63

## Write bibliography information
bib <- c(
    sgejobs = citation('sgejobs'),
    BiocStyle = citation('BiocStyle'),
    glue = citation('glue'),
    knitcitations = citation('knitcitations'),
    knitr = citation('knitr')[3],
    lubridate = citation('lubridate'),
    pryr = citation('pryr'),
    purrr = citation('purrr'),
    R = citation(),
    remotes = citation('remotes'),
    rmarkdown = citation('rmarkdown')[1],
    sessioninfo = citation('sessioninfo'),
    testthat = citation('testthat'),
    tidyr = citation('tidyr')
)

write.bibtex(bib, file = 'basics.bib')
```


# Basics

## Install `r Githubpkg('LieberInstitute/sgejobs')`

`R` is an open-source statistical environment which can be easily modified to enhance its functionality via packages. `r Githubpkg('LieberInstitute/sgejobs')` is a `R` package available via GitHub. `R` can be installed on any operating system from [CRAN](https://cran.r-project.org/) after which you can install `r Githubpkg('LieberInstitute/sgejobs')` by using the following commands in your `R` session:

```{r 'installpkg', eval = FALSE}
if (!requireNamespace("remotes", quietly = TRUE))
   install.packages("remotes")

remotes::install_github('LieberInstitute/sgejobs')
```


## Asking for help

As package developers, we try to explain clearly how to use our packages and in which order to use the functions. But `R` and `JHPCE/SGE` have a steep learning curve so it is critical to learn where to ask for help. For JHPCE questions, please use check the [JHPCE help page](https://jhpce.jhu.edu/contact/). For `r Githubpkg('LieberInstitute/sgejobs')` please post issues in GitHub. However, please note that if you want to receive help you should adhere to the [posting guidelines](http://www.bioconductor.org/help/support/posting-guide/). It is particularly critical that you provide a small reproducible example and your session information so package developers can track down the source of the error.

## Citing `r Githubpkg('LieberInstitute/sgejobs')`

We hope that `r Githubpkg('LieberInstitute/sgejobs')` will be useful for your research. Please use the following information to cite the package and the overall approach. Thank you!

```{r 'citation'}
## Citation info
citation('sgejobs')
```

# Overview

`r Githubpkg('LieberInstitute/sgejobs')` is TODO

```{r 'intro'}
library('sgejobs')
```


# Reproducibility

The `r Githubpkg('LieberInstitute/sgejobs')` package `r citep(bib[['sgejobs']])` was made possible thanks to:

* R `r citep(bib[['R']])`
* `r Biocpkg('BiocStyle')` `r citep(bib[['BiocStyle']])`
* `r CRANpkg('glue')` `r citep(bib[['glue']])`
* `r CRANpkg('knitcitations')` `r citep(bib[['knitcitations']])`
* `r CRANpkg('lubridate')` `r citep(bib[['lubridate']])`
* `r CRANpkg('pryr')` `r citep(bib[['pryr']])`
* `r CRANpkg('purrr')` `r citep(bib[['purrr']])`
* `r CRANpkg('remotes')` `r citep(bib[['remotes']])`
* `r CRANpkg('rmarkdown')` `r citep(bib[['rmarkdown']])`
* `r CRANpkg('sessioninfo')` `r citep(bib[['sessioninfo']])`
* `r CRANpkg('testthat')` `r citep(bib[['testthat']])`
* `r CRANpkg('tidyr')` `r citep(bib[['tidyr']])`


Code for creating the vignette

```{r createVignette, eval=FALSE}
## Create the vignette
library('rmarkdown')
system.time(render('sgejobs-quickstart.Rmd', 'BiocStyle::html_document'))

## Extract the R code
library('knitr')
knit('sgejobs-quickstart.Rmd', tangle = TRUE)
```

Date the vignette was generated.

```{r reproduce1, echo=FALSE}
## Date the vignette was generated
Sys.time()
```

Wallclock time spent generating the vignette.

```{r reproduce2, echo=FALSE}
## Processing time in seconds
totalTime <- diff(c(startTime, Sys.time()))
round(totalTime, digits=3)
```

`R` session information.

```{r reproduce3, echo=FALSE}
## Session info
library('sessioninfo')
options(width = 120)
session_info()
```

# Bibliography

This vignette was generated using `r Biocpkg('BiocStyle')` `r citep(bib[['BiocStyle']])`
with `r CRANpkg('knitr')` `r citep(bib[['knitr']])` and `r CRANpkg('rmarkdown')` `r citep(bib[['rmarkdown']])` running behind the scenes.

Citations made with `r CRANpkg('knitcitations')` `r citep(bib[['knitcitations']])`.

```{r vignetteBiblio, results = 'asis', echo = FALSE, warning = FALSE, message = FALSE}
## Print bibliography
bibliography()
```

Download the [bibliography](basics.bib) file.
